---
title: "Higher order derivatives, partial derivatives, plotting functions in ggplot"
output: html_document
date: "2023-08-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
install.packages("palmerpenguins")
library(palmerpenguins)
```
**Plotting a function in ggplot2**
```{r}
eq = function(x){3*x^2 + 4} # Define function

# Plot it over a specific range using the `geom_function` mark: 
ggplot(data.frame(x = c(1, 100)), aes(x = x)) + 
  geom_function(fun = eq)
#ggplot(data.frame(x = c(1, 100)), aes(x = x)): This initializes a basic ggplot plot using a data frame with a single variable x that takes values from 1 to 100. The aes(x = x) specifies that the variable x will be used on the x-axis of the plot.

#geom_function(fun = eq): This adds a geometric layer to the plot using the geom_function function. The fun parameter specifies the mathematical function that you want to plot. In this case, it uses a function called eq.
```

```{r}
eq_2 = function(x){2.4 - 5*x^3 + 2.1*x^2} # Define function

ggplot(data.frame(x = c(-50,50)), aes(x = x)) + 
  geom_function(fun = eq_2)
```
**Plotting Derivatives**
```{r}
# Store the function C(t)
ct <- function(t) {t ^ 3}

# Plot it:
ggplot(data.frame(t=c(-5, 5)), aes(x=t)) + 
  geom_function(fun=ct)
```
```{r}
# Find the derivative:
dc_dt <- D(expression(t^3), 't')
dc_dt
```
dc_dt <- D(expression(t^3), 't'): This line of code calculates the derivative of the mathematical expression t^3 with respect to the variable t.

expression(t^3) defines the mathematical expression t^3.
D(..., 't') is a function that computes the derivative of an expression with respect to a specified variable 't'.
```{r}
# Store the derivative as a function:
dc_dt_fun <- function(t) {3 * t ^2}

# Then plot them together! 
ggplot(data.frame(t = c(-5, 5)), aes(x = t)) + 
  geom_function(fun = ct, color = "red") +
  geom_function(fun = dc_dt_fun, color = "blue")
```
**Partial Derivatives**
Specify the variable you want to take the derivative with respect to.

f(x,y,z)=2xy−3x2z3

Find all partials, ∂f/∂x, ∂f/∂y and ∂f/∂z
```{r}
f_xyz <- expression(2*x*y - 3*(x^2)*(z^3))

# Partial with respect to x: 
df_dx <- D(f_xyz, 'x')
df_dx
```
```{r}
# Partial with respect to y: 
df_dy <- D(f_xyz, 'y')
df_dy
```
```{r}
# Partial with respect to z: 
df_dz <- D(f_xyz, 'z')
df_dz
```
Find partial derivatives by defining the expression, then feeding the expression and partial into D(expression, partial derivative variable)

**Final example**
Given the function:
P(q)=2cos(3q+0.4)−5.6e1.4q

Find the instantaneous slope at q=0.8
```{r}
# First, create the expression: 
pq <- expression(2 * cos(3 * q + 0.4) - 5.6 * exp(1.4 * q))

# Find the first derivative with respect to q:
dp_dq <- D(pq, 'q')

# Return the first derivative:
dp_dq
```
```{r}
# Define value of q
q <- 0.8

# Evaluate dp_dq at that value: 
eval(dp_dq)
```
**ggplot**
Customize graph using penguins dataset from palmerpenguins 
- Adding axis labels, titles & subtitles
- Switching x & y axis
- A few new graph types
- Faceting
- Updating colors
```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
         geom_point(aes(color = species)) +
  scale_color_manual(values = c("purple","cyan3","goldenrod")) +
  labs(x = "Flipper length (mm)",
       y = "Body mass (g)",
       title = "Palmer penguin size measurements",
       subtitle = "Palmer Archipelago, Antarctica (2007 - 2009)",
       caption = "Collected by Dr. Kristen Gorman and colleages at Palmer Station LTER") +
  coord_flip() +
  facet_wrap(~island) +
  theme_minimal()
```
- ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)): This sets up the base plot using the penguins dataset. It maps the flipper_length_mm column to the x-axis and the body_mass_g column to the y-axis using the aes() (aesthetic) function.

- geom_point(aes(color = species)): This adds a scatter plot layer to the base plot. Each point represents a penguin's flipper length and body mass. The color = species mapping assigns a color to each point based on the penguin's species.

- scale_color_manual(values = c("purple", "cyan3", "goldenrod")): This customizes the color scale used for the species aesthetic. It assigns specific colors to each species: "purple" for Adelie penguins, "cyan3" for Chinstrap penguins, and "goldenrod" for Gentoo penguins.

- labs(...): This adds labels and titles to the plot. It sets the x-axis label, y-axis label, plot title, subtitle, and caption.

- coord_flip(): This flips the x and y axes, effectively transposing the plot. This is often used to create horizontal bar-like plots when working with categorical variables.

- facet_wrap(~island): This creates separate subplots for different levels of the island variable. Each subplot displays the scatter plot for a specific island in the Palmer Archipelago.

- theme_minimal(): This applies a minimalistic theme to the plot, changing the overall appearance of the plot elements such as axis lines, labels, and grid lines.