---
title: "EDS212 Session 1B"
output: html_document
date: '2023-08-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("vctrs")
library(vctrs)
install.packages("tidyverse", dependencies = TRUE)
library(tidyverse)
```
Exponents and logs in R
- log()=natural log
- log10()=log base 10
- exp()=natual exponential 

To make sequence of values stored in a vector in R, use the seq() function
- seq(from = start_value, to = end_value, by = increment)
- To create a sequence from 2-18 by increments of 0.3, use:

```{r seq}
seq(from = 2, to = 18, by = 0.3)
```
Use the length= argument within seq() to specify amount of values you want

To get 30 values between 2 and 18, evenly spaced, use: 
```{r}
seq(from = 2, to = 18, length = 30)
```
- Use keyboard shortcut "cmd+opt+I" to make a new code chunk
- Enclosing text in ** or __ on either end makes it **bold**; Enclosing text in * or _ on either end makes it _italic_; enclosing text in **_ or __* on either end makes it **_bold and italic_**

**Making the logistic growth function**
- Use the shortcut "Cmd+opt+X"
- Expression for population size at anytime following logistic growth is
  - Nt=K/1+(K-N0/N0)e^-rt
```{r}
pop_logistic <- function(capacity, init_pop, rate, time_yr) {
  capacity / (1 + ((capacity - init_pop) / init_pop) * exp(-rate * time_yr))
}
```

Predict the estimated population of chipmunks in one region one time
- K=2580 individuals
- r= 0.32(yr^-1)
- N0=230 individuals
- t=2.4 years
```{r}
pop_logistic(capacity = 2580, init_pop = 230, rate = 0.32, time_yr = 2.4)
```
Predict the estimated population of chipmunks in one region many times using a sequence for the time input
- 0-20 years by 1/2 year increments 
```{r}
# First, create the vector (a sequence of values)
time_vec <- seq(from = 0, to = 20, by = 0.5)

# Then, use that as your time input in the model and save the output in a variable:
chipmunk_pop<- pop_logistic(capacity = 2580, init_pop = 230, rate = 0.32, time_yr = time_vec)
```
**Make a Plot**
Create a dataframe from out time sequence and the predicted populations
```{r}
chipmunk_df <- data.frame(time_vec, chipmunk_pop)

head(chipmunk_df)

ggplot(data = chipmunk_df, aes(x = time_vec, y = chipmunk_pop)) +
  geom_point()
```

